---
title: "Bar Plots with Error Bars"
format: 
   html:
    toc: true
    code-fold: true
    code-summary: "Show the code"
editor: visual
---

### References:

-   [Rajendra Choure - Grouped bar plot with error bars](https://www.youtube.com/watch?v=-7rpZ_yYpEE&t=11s)\
-   [Rajendra Choure - Adding data labels](https://www.youtube.com/watch?v=WfeqvyId2sU)

Next: Check out [*Empiracle Cumulative Distribution Function* (ECDF)](https://www.youtube.com/watch?v=iI8RA6fjxNg)

```{r}
#| echo: false 
#| message: false
#| warning: false 
#| 
library(tidyverse) 
library(ggsci) # for color palettes of scieintifc journals
library(ggpubr) # for theme pubr
library(ggplot2)
library(plotly)
# library(dplyr)
library(tidyr)

```

```{r}

iris_plot =  gather(iris,FlPrt,Length,1:4)

  iris_plot= group_by(iris_plot, Species,FlPrt)

  iris_plot=summarise(iris_plot,mnL=mean(Length),sdL=sd(Length)) 

iris_plot=  ggplot(iris_plot,aes(reorder(Species,mnL),mnL,fill=reorder(FlPrt,mnL)))+
  geom_bar(stat="identity",position="dodge")+
  geom_text(mapping=aes(label=mnL),position=position_dodge(width=0.9),
            cex=2.5,vjust=-4)+
  labs(title="Iris: Measurment of floral organs",
       subtitle="Length abd width of srpal abd petal",
       x="Species",y="Length ib cm",
       caption="Data: iris dataset by Edgar Anderson",fill="Description")+
  geom_errorbar(mapping=aes(ymin=mnL-sdL,ymax=mnL+sdL),
                width=0.2,position=position_dodge(width=0.9))+
  theme_classic()+scale_fill_jco()
```

```{r}

iris_plot <- iris %>% 
  gather(FlPrt, Length, 1:4) %>% 
  group_by(Species, FlPrt) %>% 
  summarise(mnL = mean(Length), sdL = sd(Length)) %>% 
  
  ggplot(aes(Species, mnL, fill = FlPrt)) +
  
  # identity is required to use values for height of bar,
  # otherwise, Count is used
  
  # Dodge is used to group bar plot, without it you get stacked bar
  geom_bar(stat = "identity", position = "dodge") 
   
iris_plot
```

### Reorder categories

-   Setosa \> versicolor \> verginica

-   Reorder(categorical Variable, numeric variable)

-   reorder(Species, mnL)

```{r}
#| fig-cap: "Ordered bar plot"
iris %>% 
  gather(FlPrt, Length, 1:4) %>% 
  group_by(Species, FlPrt) %>% 
  summarise(mnL = mean(Length), sdL = sd(Length)) %>% 
  
  # reorder
  #  ggplot(aes(Species, mnL, fill = FlPrt))  
  # first argument is category, second argument is the numeric values
  ggplot(aes(reorder(Species, mnL), mnL, fill=reorder(FlPrt, mnL))) +
  
  # identity is required to use values for height of bar,
  # otherwise, Count is used
  
  # Dodge is used to group bar plot, without it you get stacked bar
  geom_bar(stat = "identity", position = "dodge") 
   
iris_plot
```

### Add data values to bar

-   map the value to the label attribute\

-   geom_text()

-   geom_text(mapping aes(label=mnL), postition = position_dodge(width = 0.9), cex = 2.5, vjust(= -4) )

```{r}
#| fig-cap: "Ordered bar plot with geom_text() "
iris %>% 
  gather(FlPrt, Length, 1:4) %>% 
  group_by(Species, FlPrt) %>% 
  summarise(mnL = mean(Length), sdL = sd(Length)) %>% 
  
  # reorder
  #  ggplot(aes(Species, mnL, fill = FlPrt))  
  # first argument is category, second argument is the numeric values
  ggplot(aes(reorder(Species, mnL), mnL, fill=reorder(FlPrt, mnL))) +
  
  # identity is required to use values for height of bar,
  # otherwise, Count is used
  
  # Dodge is used to group bar plot, without it you get stacked bar
  geom_bar(stat = "identity", position = "dodge") +
  
  geom_text(mapping = aes(label = mnL), 
            position = position_dodge(width = 0.9),
            cex = 2.5, vjust =  -2.5)
   
```

### Add titles to plot elements

```{r}
#| fig-cap: "Ordered bar plot with geom_text() "
iris %>% 
  gather(FlPrt, Length, 1:4) %>% 
  group_by(Species, FlPrt) %>% 
  summarise(mnL = mean(Length), sdL = sd(Length)) %>% 
    ggplot(aes(reorder(Species, mnL), mnL, fill=reorder(FlPrt, mnL))) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(mapping = aes(label = mnL), 
            position = position_dodge(width = 0.9),
            cex = 2.5, vjust =  -2.5) + 
  labs(title = "Iris: Measurement of floral organs",
       subtitle = "Length and width of sepal and petal",
       x = "Species", y = "Length in cm",
       caption = "Data: iris dataset by Edger Anderson", 
       fill = "Description")

```

### Add error bars to illustrate uncertainty

-   Add error bars with *geom_errorbar()*\

```{r eval = FALSE, echo=TRUE}
geom_errorbar(mapping = aes(ymin=mnL-sdL, 
                              ymax = mnL +sdL),
                              width = 0.2, 
                              position = position_dodge(width = 0.9))

```

```{r}
iris %>% 
  gather(FlPrt, Length, 1:4) %>% 
  group_by(Species, FlPrt) %>% 
  summarise(mnL = mean(Length), sdL = sd(Length)) %>% 
    ggplot(aes(reorder(Species, mnL), mnL, fill=reorder(FlPrt, mnL))) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(mapping = aes(label = mnL), 
            position = position_dodge(width = 0.9),
            cex = 2.5, vjust =  -7) + 
  labs(title = "Iris: Measurement of floral organs",
       subtitle = "Length and width of sepal and petal",
       x = "Species", y = "Length in cm",
       caption = "Data: iris dataset by Edger Anderson", 
       fill = "Description") + 
  
  geom_errorbar(mapping = aes(ymin = mnL - sdL,
                              ymax = mnL + sdL),
                width = 0.2,
                position = position_dodge(width = 0.9))
  
```

### Add theme *ggpubr* and color palettes *ggsci*.

-   *library(ggpubr)*
-   *library(ggsci)*

```{r color-palettes}

library(ggpubr)
library(ggsci)
iris %>% 
  gather(FlPrt, Length, 1:4) %>% 
  group_by(Species, FlPrt) %>% 
  summarise(mnL = mean(Length), sdL = sd(Length)) %>% 
    ggplot(aes(reorder(Species, mnL), mnL, fill=reorder(FlPrt, mnL))) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(mapping = aes(label = mnL), 
            position = position_dodge(width = 0.9),
            cex = 2.5, vjust =  -7) + 
  labs(title = "Iris: Measurement of floral organs",
       subtitle = "Length and width of sepal and petal",
       x = "Species", y = "Length in cm",
       caption = "Data: iris dataset by Edger Anderson", 
       fill = "Description") + 
  
  geom_errorbar(mapping = aes(ymin = mnL - sdL,
                              ymax = mnL + sdL),
                width = 0.2,
                position = position_dodge(width = 0.9)) +
  
  # themes 
  # theme_bw()
  # theme_classic()  
  
  # ggpubr::theme_pubr()  
  
  # color palettes  
 
  
  theme_pubr() +
#  scale_fill_jco() 
  # scale_fill_npg()  # ***
  # scale_fill_aaas()
  # scale_fill_jama()  # ***  
#  scale_fill_lancet() # *** 
#  scale_fill_ucscgb() # :( 
# scale_fill_d3()
# scale_fill_locuszoom() # :)  
# scale_fill_igv() # :)  
# scale_fill_startrek() # :)  
# scale_fill_futurama() # :)  
# scale_fill_rickandmorty() # :)  
scale_fill_simpsons() # :)


```
