---
title: "Ordering Bar Plots with *reorder()*"
date: "2022-06-11"
format: 
   html:
    toc: true
    code-fold: true
    code-summary: "Show the code"
editor: visual
--- 

```{r}
#| echo: false
#| warning: false
#| message: false 
#| 
library(tidyverse) 
library(kableExtra)
library(ggplot2)

```


**References**: 

* [Bargraph in R , Arrange bars of barplot using R and GGplot by Rajendra Choure](https://www.youtube.com/watch?v=nQDURKIF4og&list=PLfAzV0jqypOhtjvLvYyMG-6_viNCKLLXP&index=4)  


### Create data using rnorm()  

* rorm() syntax:  *rnorm(n, mean, sd)*
```{r}
subject <- rep(LETTERS[1:5], each = 20)

# rnorm syntax rnorm(n, mean, sd)
scores <- c(rnorm(20, 45, 5), rnorm(20, 35, 9), rnorm(20, 45, 3), rnorm(20, 28,5), rnorm(20,31,8)) 

dfs <- scores
df <- data.frame(subject, scores = round(scores,0))

head(df) %>% 
 
  kbl(caption = "Subject Scores") %>% 
  kable_classic(full_width = F, html_font = "Cambria", position = "left")  
  
```

```{r}

df_sm <- df %>% 
  group_by(subject) %>% 
  summarise(mean = mean(scores), sd = sd(scores)) 

df_sm
```

<br>  

### Plot values using *geom_bar(stat= "identity")*    
```{r}
ggplot(df_sm, aes(subject, mean, fill = subject)) +
  geom_bar(stat = "identity")
```



- *geom_bar(stat = "identity")* uses the variable and not the count of the variable.  

- *sd* is included to plot error bars  
```{r eval = FALSE}

# Question:  Can you use geom_bar for counts ? 
# Qusetion:  How do you use stat_summary() ? 
# Question:  How wouldl you show bars as proportion, instead of counts or values?  
ggplot(df_sm, aes(subject, mean, fill = subject)) +
 # geom_bar(stat_count())
  geom_bar(stat = scores)
geom_bar(inherit.aes = subject)
```  

<br> 

### Ordering the data makes it easier to see which category is best and worse.  

* Modify *ggplot(df_sm, aes(subject, mean, fill = subject))* 
* Reorder by *mean*: *aes(reorder(subject, mean), mean, fill = subject))*
```{r}

ggplot(df_sm, 
       # reoder by mean 
       aes(reorder(subject, mean), mean, fill = subject)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = mean-sd, ymax = mean + sd), width = 0.2) +
  theme_classic() + 
  theme(legend.position = "none")
```


To change the order, place -mean
```{r}

ggplot(df_sm, 
       # reoder by mean 
       aes(reorder(subject, -mean), mean, fill = subject)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = mean-sd, ymax = mean + sd), width = 0.2) +
  theme_classic() + 
  theme(legend.position = "none")
```

